Supplementary information: Variational theory of non-relativistic quantum electrodynamics

### 2019-01-27 supp2.tex
slight edit after johannes' comment

\textcolor{blue}{It is also of interest to consider "field-only" observables such as the number of (virtual) photons in the ground state. Calculating the wavefunction of the ground state at first-order in perturbation theory, and calculating the probability of a photon being in the ground state, one finds that 
\begin{equation}
\langle N \rangle = \frac{q^2}{m^2} \sum\limits_{b\neq a}\sum\limits_{n=1}^{N_p} \frac{1}{2\hbar\epsilon_0L}\frac{|p_{ba}|^2\sin^2\left(\frac{n\pi z}{L}\right)}{\omega_n(\omega_a-\omega_b-\omega_n)^2}.
\end{equation}
In Fig. S1, we compare the expectation value of the number of photons calculated variationally, perturbatively, and with exact diagonalization. Unlike all other cases shown in this work (energy and $\langle A\cdot p \rangle$, the number of photons does not agree between variational and exact diagonalization. However, this behavior does not reflect any inconsistency with the variational method, but is instead a feature which is to be expected. In particular, the physics of the ansatz of Eq. 4 of the main text is that a complicated ground state which has many virtual photons (see numerical diagonalization result) in it should be physically equivalent to a ground state with nearly no virtual photons (see variational result), but with different modes and frequencies. In other words, the reason the number of virtual photons differs is simply because they are being counted in the basis of "interacting modes" shown in Fig. 2b of the main text.   That said, our variational theory does not give a prescription to translate the interacting mode operators to the bare mode operators.}

\begin{figure}[t]
\includegraphics[width=6cm]{FigureS1.pdf}
\caption{\textcolor{blue}{\textbf{Number of virtual photons in the ground state calculated variationally, numerically, and through perturbation theory.} Parameters are the same as in Fig. 2 (top panel) of the main text.}}
\label{fig:ansatz}
\end{figure}

>>>>

\textcolor{blue}{It is also of interest to consider "field-only" observables such as the number of (virtual) photons in the ground state. Calculating the wavefunction of the ground state at first-order in perturbation theory, and calculating the probability of a photon being in the ground state, one finds that 
\begin{equation}
\langle N \rangle =\sum\limits_{n=1}^{N_p} \langle a^{\dagger}_{n,\text{bare}}a_{n,\text{bare}} \rangle = \frac{q^2}{m^2} \sum\limits_{b\neq a}\sum\limits_{n=1}^{N_p} \frac{1}{2\hbar\epsilon_0L}\frac{|p_{ba}|^2\sin^2\left(\frac{n\pi z}{L}\right)}{\omega_n(\omega_a-\omega_b-\omega_n)^2},
\end{equation}
where the "b" or "bare" subscript in the creation and annihilation operators denotes that these are creation and annihilation operators for photon modes in the cavity in the absence of interactions (e.g., the sine modes for the 1D cavity considered here). In Fig. S1, we compare the expectation value of the number of "bare" photons calculated  perturbatively, and with exact diagonalization with the variationally calculated number of "interacting" photons, $\sum\limits_{n=1}^{N_p} \langle a^{\dagger}_{n,\text{int}}a_{n,\text{int}} \rangle$ (e.g., the modes shown in Fig. 2b of the main text). The number of photons in the bare basis differs substantially from the number of photons in the interacting basis.}

\textcolor{blue}{This behavior does not reflect any inconsistency with the variational method, but is instead a feature which is to be expected. In particular, the physics of the ansatz of Eq. 4 of the main text is that a complicated ground state which has many virtual photons (see numerical diagonalization result) in it should be physically equivalent to a ground state with nearly no virtual photons (see variational result), but with different modes and frequencies. In other words, the reason the number of virtual photons differs is simply because they are being counted in the basis of "interacting modes" shown in Fig. 2b of the main text.   That said, our variational theory does not give a prescription to translate the interacting mode operators to the bare mode operators.}

\begin{figure}[t]
\includegraphics[width=6cm]{FigureS1.pdf}
\caption{\textcolor{blue}{\textbf{Number of virtual photons (bare and interacting) in the ground state calculated variationally, numerically, and through perturbation theory.} Parameters are the same as in Fig. 2 (top panel) of the main text.}}
\label{fig:ansatz}
\end{figure}
